
---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CompMod

<!-- badges: start -->
<!-- badges: end -->

`CompMod` provides tools for performing *comparable company analysis (comps)*, a core valuation methodology in investment banking and equity research. The package covers the full workflow: reading standardised valuation data from Excel, computing key trading multiples, and summarising peer group statistics to support an implied valuation range.

## Installation

You can install the development version of `CompMod` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("yourgithub/CompMod")
```

## Examples

```{r setup}
library(CompMod)
```

#### Compute trading multiples

```{r}
comps <- calc_multiples(valuation_data)
comps[, c("Company.Name", "EV.Revenue", "EV.EBITDA", "P.E")]
```

#### Summarise the peer group

```{r}
peer_summary(comps)
```
