---
title: "Introduction to CompMod"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to CompMod}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Motivation

Commonly called "comps," Comparable company analysis is one of the most widely used valuation methodologies in investment banking and equity research. The idea is straightforward: if you want to know what a private company or an acquisition target is worth, look at how the market is pricing similar public companies and apply those multiples to your target.

In practice, however, building a comps model involves several repetitive steps: importing data from Excel, re-deriving multiples to verify them, and computing peer group statistics (mean, median, min, max) to set a valuation range. These steps are tedious, error-prone, and hard to reproduce.

`CompMod` solves exactly this problem. It provides three focused functions that turn a raw Excel file of peer company data into a clean, reproducible comps analysis in just a few lines of R.

## Installation

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("yourgithub/CompMod")
```

Load the package:

```{r setup}
library(CompMod)
```

## The `valuation_data` dataset

`CompMod` ships with a built-in dataset of 15 fictional technology companies, each with the financial metrics needed to run a comps analysis.

```{r}
head(valuation_data[, c("Company.Name", "Ticker", "Share.Price",
                         "Enterprise.Value", "Revenue..LTM.", "EBITDA")])
```

The dataset contains `r nrow(valuation_data)` companies and `r ncol(valuation_data)` columns covering share price, capital structure metrics, income statement items, and pre-computed multiples.

## Step 1 — Read your Excel file with `read_valuation_excel()`

In a real workflow you would load your own Excel file. `read_valuation_excel()` reads the file, standardizes column names, and validates that all required columns and numeric fields are present before any analysis runs.

```{r, eval = FALSE}
df <- read_valuation_excel("path/to/your/valuation_data.xlsx")
```

If the file is missing required columns, the function stops immediately with a messages 

## Step 2 — Compute multiples with `calc_multiples()`

`calc_multiples()` takes the data frame and computes five trading multiples from first principles:

| Multiple | Formula |
|---|---|
| EV/Revenue | Enterprise Value divided by LTM Revenue |
| EV/EBITDA | Enterprise Value divided by EBITDA |
| P/E | Share Price divided by EPS |
| Price/Sales | Market Cap divided by LTM Revenue |
| EBITDA Margin | EBITDA divided by LTM Revenue |

```{r}
comps <- calc_multiples(valuation_data)
comps[, c("Company.Name", "EV.Revenue", "EV.EBITDA", "P.E", "EBITDA.Margin")]
```

## Step 3 — Summarise the peer group with `peer_summary()`

The final step condenses the peer group multiples into a summary table of mean, median, min, and max. This is what an analyst uses to derive an implied valuation range for a target company.

```{r}
peer_summary(comps)
```

You can also pass a custom subset of multiples:

```{r}
peer_summary(comps, multiples = c("EV.Revenue", "EV.EBITDA"))
```

## Putting it all together

Here is the full workflow from raw Excel to peer group summary in three lines:

```{r, eval = FALSE}
library(CompMod)

df    <- read_valuation_excel("valuation_data.xlsx")
comps <- calc_multiples(df)
peer_summary(comps)
```

## Applying the results

Once you have the peer group median multiples, you can apply them to a target company's financials. For example, if the peer group median EV/EBITDA is *12.0x* and your target's EBITDA is *$500M*:

- Implied Enterprise Value = 12.0 x $500M = **$6,000M**
- Subtract Net Debt to get Implied Equity Value
- Divide by Shares Outstanding to get Implied Share Price

`CompMod` gives you the summary statistics — the final step of applying them to a specific target is left to the analyst to ensure business context is properly considered.
